---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)


```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}

# Insert your code here, create more chunks as necessary
if(!(c("psych") %in% rownames(installed.packages()))) {install.packages('psych')}
library(psych)

if(!(c("pastecs") %in% rownames(installed.packages()))) {install.packages('pastecs')}
library(pastecs)

stat.desc(inc) 
describe(inc)
tail(inc)
table(inc$Industry)
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}

# Answer Question 1 here
if(!(c("ggplot2") %in% rownames(installed.packages()))) {install.packages('ggplot2')}
library(ggplot2)

ggp<-ggplot(data=inc, aes(x=State)) 
# counts
ggp + geom_histogram(fill="black",stat = "count") + coord_flip()+
theme_classic()+
    labs(title="Distribution of Companies", x="State",y="Number of Companies")

```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


Let's rank the states by the number of companies they have, and find the 3rd state with most employees
```{r}
if(!(c("dplyr") %in% rownames(installed.packages()))) {install.packages('dplyr')}
library(dplyr)

df <- group_by(inc, State)
df <-summarise(df,count = n())
df <- arrange(df, desc(count))

#state with 3rd most companies is NY
df$State[3]
```

We can see that the state with 3rd most companies is NY


Let's filter the data for only NY state records and return only the records with no missing values

```{r}
NY_inc <- filter(inc, State == df$State[3])
summary(NY_inc)
NY_inc <- NY_inc[complete.cases(NY_inc),]

```
Let's summarize the data and identify outliers
outliers are those observations that lie outside 1.5*IQR, where IQR, the 'Inter Quartile Range' is the difference between 75th and 25th quartiles
```{r}
# aggregate employees by Industry
Employment_by_Industry<- group_by(NY_inc,Industry) %>% 
  summarise(Avg_employee_cnt = mean(Employees), Median_Employee_cnt = median(Employees), Total_employee_cnt = sum(Employees) 
            , Min_employee_cnt = min (Employees), Max_employee_cnt = max(Employees) , 
    high_Outlier_limit = quantile (Employees)[4] + 1.5*IQR(Employees), low_Outlier_limit = quantile (Employees)[2] - 1.5*IQR(Employees)
            ,First_quartile = quantile(Employees)[2],
            First_quartile = quantile(Employees)[2] , third_quartile = quantile (Employees)[4])
Employment_by_Industry
```


Let's create a plot that shows the average and median employment by industry for companies in NY state, without dealing with outliers and showing variability using boxplots. Points outside the whiskers of the boxplot are outliers

```{r}


g <- ggplot(NY_inc, aes(x = Industry, y = Employees)) +
        geom_boxplot()
  g+ coord_flip() 
  

```

Let's remove the outliers and reproduce the boxplot

```{r}
g <- ggplot(NY_inc, aes(x = Industry, y = Employees)) +
        geom_boxplot(outlier.shape = NA)
  g + coord_flip() +
  scale_y_continuous(limits = quantile(NY_inc$Employees, c(0.1, 0.9)))
```


## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

Let's get only the cases without missing values
```{r}

# Answer Question 3 here
INC_2 <- inc[complete.cases(inc),]

INC_summary <- group_by(INC_2,Industry) %>% 
  summarise( Total_employee_cnt = sum(Employees) 
            , Total_Revenue = sum (Revenue), Revenue_per_employee = sum (Revenue)/sum(Employees))

INC_summary2 <-arrange(INC_summary, desc(Revenue_per_employee))
INC_summary2
ggp<- ggplot(data=INC_summary2, aes(x=Industry, y=Revenue_per_employee)) 
# counts
ggp + geom_bar(fill="black",stat = "identity") + coord_flip()

```


